---
title: "Analysing the adoption and Growth of Bitcoin in relation to scarcity"
name: "Afnan UR Rehman"
Student ID: "21324930"
output: html_notebook
---


*Choosing the CSV File:*

```{r}
sfdata <- read.csv(choose.files())
```

```{r}
install.packages("lmtest")

```


```{r}
library(car)
library(tseries)
library(sandwich)# needed to produce heteroskedastic consistent standard errors
library(lmtest) # needed for Breusch Pagan Test - alternative to White's test

```


```{r}
attach(sfdata)
```

```{r}
summary(sfdata)
```


* Ploting a simple relationship between vairables with price as a dependent variable and stockflow and market-cap as independent vairables: *

```{r}
plot(price ~ stockflow, data = sfdata)
```
* Converting the values into log in order to remove the inconsistencies: *

```{r}
sfdata$log_price <- log(sfdata$price)
sfdata$log_stockflow <- log(sfdata$stockflow)
sfdata$log_active_addresses<-log(sfdata$active_addresses)
sfdata$log_blocks <- log(sfdata$blocks)
sfdata$log_market_cap <- log(sfdata$market_cap)
```

* Ploting after conversion to see the relationship between variables: *

```{r}
plot(log_price ~ log_stockflow, data = sfdata)
```


```{r}
plot(log_price ~ log_market_cap, data = sfdata)
```




* Plotting to test the Hypothesis using SIMPLE LINEAR REGRESSION: *


*Fitting the linear regression model to predict the behavior of price and stock flow ratio to test the hypothesis 1:*
```{r}
lnmodel<- lm(log_price ~ log_stockflow, data = sfdata)
summary(lnmodel)
```

* Plotting the linear regression model 1 *


```{r}
plot(log_price ~ log_stockflow, data = sfdata)
abline(lnmodel, col='red') # fitted values
```




* Compute the OLS coefficients and their corresponding standard errors *

```{r}
Intercept <- rep(1, nrow(sfdata))
# Define the matrix X by joining the intercept term with the independent variable i.e, log_stockflow
X <- cbind(Intercept, sfdata$log_stockflow)
# Multiply the transpose of X to itself
XtX <- t(X) %*% X
# Compute the inverse of X
XtX_inv <- solve(XtX)
# Define the vector Y which denotes the dependent variable - log_price
Y <- sfdata$log_price
# Multiply the transpose of X with Y
XtY <- t(X) %*% Y
# Compute the estimates
beta <- XtX_inv %*% XtY
# Sigma represents the variance of the residuals
sigma <- var(lnmodel$residuals)
```

```{r}
var_beta <- sigma * XtX_inv
paste("The estimated coefficient is", round(beta[2],4))
```



```{r}
paste("The standard error of the above coefficient is",round(sqrt(var_beta[4]),4))
```
```{r}
paste("The t-statistic for the above coefficient is",round(beta[2]/sqrt(var_beta[4]),4))
```

```{r}
confint(lnmodel, "log_stockflow")
```

*Since Market cap is a function of price calculated as
Market cap = price x circulating supply,
market cap is dependent vairable but since it is a funtion of price, it is evident that it will form a perfect linear regression model with very high r-square value which can create model biasness and which in turn could inflate the vairance of coefficients estimates and make the model unstable.* 

*However, we can form a Multiregression model in order to study the relationship between stock_flow, price, active addressess and the market cap and prove a joint hypothesis i.e, what are the other factors along with the scarcity of bitcoin that are affecting the price of bitcoin keeping the price of bitcoin as a dependent variable consistently to maintain a stable model. We might exclude the marketcap from the joint hypothesis in order to maintain the stability of the model*



```{r}
Multilnmodel <- lm(log_price ~ log_stockflow + log_active_addresses + log_market_cap , data = sfdata)
summary(Multilnmodel)
```


```{r}
hist(Multilnmodel$residuals, breaks = 30, freq = FALSE,
main = "Histogram of Residuals with Normal Distibution Overlay")
lines(seq(-1, 0.5, by=.05), dnorm(seq(-1, 0.5, by = .05), mean(Multilnmodel$residuals),
sd(Multilnmodel$residuals)), col="blue")
```

```{r}
qqPlot(Multilnmodel$residuals)
```

```{r}
# Jacque-Bera test for assessing Normality 
# The null hypothesis is that the residuals are normally distributed
jarque.bera.test(Multilnmodel$residuals)
```

* Test the joint hypothesis that the coefficients of the two variables are all zero
i.e., H0 : β1 = β2 = β3 = 0  *

```{r}
linearHypothesis(Multilnmodel, c("log_stockflow = 0","log_active_addresses = 0","log_market_cap=0"))
```



* We can test the model stability and try to improve the model by removing the market cap since market cap in itself is a function of price as discussed before and could be effecting the model predictability *

```{r}
updatedmultilnmodel <- lm(log_price ~ log_stockflow + log_active_addresses , data = sfdata)
summary(updatedmultilnmodel)
```
```{r}
hist(updatedmultilnmodel$residuals, breaks = 30, freq = FALSE,
main = "Histogram of Residuals with Normal Distibution Overlay")
lines(seq(-1, 0.5, by=.05), dnorm(seq(-1, 0.5, by = .05), mean(updatedmultilnmodel$residuals),
sd(updatedmultilnmodel$residuals)), col="blue")
```

```{r}
qqPlot(updatedmultilnmodel$residuals)
```

```{r}
# Jacque-Bera test for assessing Normality 
# The null hypothesis is that the residuals are normally distributed
jarque.bera.test(updatedmultilnmodel$residuals)
```
Test the joint hypothesis that the coefficients of the two variables are all zero
i.e., H0 : β1 = β2 = 0

```{r}
linearHypothesis(updatedmultilnmodel, c("log_stockflow = 0","log_active_addresses = 0"))
```

```{r}
mean(sfdata$log_price)
```

```{r}
mean(updatedmultilnmodel$fitted.values)
```

* Plot the fitted values from model 1 against the actual price values *

```{r}
# Plot of predicted vs actual values
plot(updatedmultilnmodel$fitted.values, sfdata$log_price,
     xlim = c(min(updatedmultilnmodel$fitted.values), max(updatedmultilnmodel$fitted.values)),
     ylim = c(min(sfdata$log_price), max(sfdata$log_price)),
     xlab = "Predicted log(price)", ylab = "Actual log(price)")
abline(0, 1, col = "red")

```

* Incorporate higher order terms of the predicted values into the specification * 
```{r}
sfdata$fits <- updatedmultilnmodel$fitted.values
sfdata$fits2 <- (sfdata$fits)^2
sfdata$fits3 <- (sfdata$fits)^3
updatedmultilnmodel_d <- lm(log_price ~ log_stockflow + log_active_addresses + fits2 + fits3, data = sfdata)
summary(updatedmultilnmodel_d)
```

* regression residuals/errors *

```{r}
mean(updatedmultilnmodel$residuals)
```



```{r}
plot(updatedmultilnmodel$fitted.values, updatedmultilnmodel$residuals)
abline(0,0)
```

* White test: *

```{r}
sfdata$residuals2 <- (updatedmultilnmodel$residuals)^2  # Correcting the calculation of squared residuals
Multilnmodel_W <- lm(residuals2 ~ log_stockflow + log_active_addresses + 
                     I(log_stockflow^2) + I(log_active_addresses^2) + 
                     log_stockflow * log_active_addresses, data = sfdata)
summary(Multilnmodel_W)

```

```{r}
RSQ <- summary(Multilnmodel_W)$r.squared
tstat <- nobs(updatedmultilnmodel)*RSQ

```

* heteroscedasticity *

```{r}
nbetas <- summary(Multilnmodel_W)$df[1]
# Compute the p value
pval <- 1 - pchisq(tstat, nbetas-1)
# An alternative to White's test - The Breusch-Pagan test
paste("The test statistic for White's test is", round(tstat,2), "with p value",
round(pval,8))

```
```{r}
bptest(updatedmultilnmodel)
```

